---
title: "Basic Data Types"
author: "Bert Gollnick"
output: 
  html_document:
    toc:true
    toc_float: true
    code_folding: hide
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Vector Creation

You can create a vector with c(). The values inside define the type of the vector.

Double variables:

```{r}
double_variables <- c(2, 3.4, 0, 7.8)
```

Integer variables:

```{r}
integer_variables <- c(2L, 4L, 15L)
```

Boolean variables:

```{r}
logical_variables <- c(TRUE, FALSE, F, T)
```

character variables:

```{r}
char_variables <- c("4", "test", 5)
```


# Types

You can check the type with typeof().

```{r}
typeof(double_variables)
```

If you want for a specific type with is.character(), is.double(), is.numeric().

```{r}
is.double(double_variables)
is.double(integer_variables)
```

is.numeric() returns true for double and integer vectors.

# Type Coercion

All elements of an atomic vector must have the same type. If you combine elements of different types they are coerced to the most flexible type.

```{r}
typeof(c(1L, 2L, 3))
```
This combination of integer and double values results in a double vector.

Logical vectors can be coerced to numeric (0 for false, 1 for true).

```{r}
as.numeric(logical_variables)
```

This is very helpful if you want to use it for filtering or aggregating.

There is an order of coercion:

logical --> integer --> numeric --> complex --> character --> list

This means you can coerce from left to right. But if you coerce from right to left you possibly lose some information.

```{r}
as.logical(as.integer(double_variables))
```

All values that are different from zero are coerced to TRUE.

```{r}
as.numeric(char_variables)
```
Due to the coercion NAs are introduced where the conversion to numeric did not produce a valid result.

