---
title: "Logistic Regression Exercise"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Understanding

We will create a Credit Approval model. More information can be found at this [link](http://archive.ics.uci.edu/ml/datasets/credit+approval).

We will predict whether an applicant received a credit.

## Data Import

Import the file to an object called "credit_approval".

```{r}
# put your code here
```

# Data Preparation

## Packages

We load required packages.

```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(keras))
suppressPackageStartupMessages(library(caret))
suppressPackageStartupMessages(library(e1071))

source("./functions/train_val_test.R")
```

In tutorial on missing data handling, this dataset was prepared and missing data was taken care of.

```{r}
# put your code here
```

The class attribute has values "-" and "+" which will be transformed to 0 and 1.

```{r}
# put your code here
```

We check that we have the same distribution as before.

```{r}
# put your code here
```

We change type of "Approved" to factor.
```{r}
# put your code here
```


## Train / Validation Split

For simplicity we will only split into training and validation

```{r}
# put your code here
```


# Modeling

```{r}
# put your code here
```

# Predictions

Please calculate predictions for a threshold of 0.5 for training and validation dataset.

```{r}
# put your code here
```

# Model Performance

## Baseline Performance 

Please calculate the accuracy, if always the most frequent class is predicted!

```{r}
# put your code here
```


## Confusion Matrix

Please calculate the confusion matrix for training and validation datset.

```{r}
# put your code here
```

```{r}
# put your code here
```


# Acknowledgement

We thank the provider of the dataset.

(confidential source)

Submitted by quinlan '@' cs.su.oz.au