---
title: "Support Vector Machines Exercise"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Understanding

We will work on heart diseases. On UCI Machine Learning Repository you find "Heart Disease" (dataset)[https://archive.ics.uci.edu/ml/datasets/Heart+Disease].

This database contains 76 attributes, but all published experiments refer to using a subset of 14 of them. In particular, the Cleveland database is the only one that has been used by ML researchers to 
this date. The "goal" field refers to the presence of heart disease in the patient. It is integer valued from 0 (no presence) to 4. Experiments with the Cleveland database have concentrated on simply attempting to distinguish presence (values 1,2,3,4) from absence (value 0). 


These are the attribute information:

1. age
2. sex
3. cp
4. trestbps
5. chol 
6. fbs
7. restecg
8. thalach
9. exang
10. oldpeak
11. slope
12. ca
13. thal
14. target

## Data Import

```{r}
# if file does not exist, download it first
file_path <- "./data/heart_disease.csv"
if (!file.exists(file_path)) {
  dir.create("./data")
  url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/heart-disease/processed.cleveland.data"
  download.file(url = url, 
                destfile = file_path)
}
```

Import the file to an object called "heart_raw".

```{r}
# place your code here
```

# Data Preparation

## Packages

We load required packages.

```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(keras))
suppressPackageStartupMessages(library(caret))
suppressPackageStartupMessages(library(e1071))

source("./functions/train_val_test.R")
```

## Column Names

Assign the column names correctly. Use these names:

age, sex, cp, trestbps, chol, fbs, restecg, thalach, exang, oldpeak,  slope, ca, thal, target


```{r}
# place your code here
```

How many levels of target-variables are present?

```{r}
# place your code here
```

Is this a classification or regression task? We will treat it as classification, but I want you to think about it.

```{r}
# Think about the task!
```

## Summary

Check the summary of the data to see if there are missing values. Are there any missing?

```{r}
# place your code here
```

## Variable Type Correction

Some numerical attributes were wrongly assigned to factors and vice versa. Check the repository description which features were wrongly assigned and correct it.

You would have to check data description. I save you some time.

- to factor:sex, cp, fbs, restecg, exang, slope, thal, target

- to numeric: ca

Don't modify the raw data. Instead create a new object "heart_mod", in which you perform the changes.

```{r}
# place your code here
```


Filter for values for "?" on "thal"-variable, because no information is available at these positions.

```{r}
# place your code here
```


## Train / Validation / Test Split

Split the data into train, validation, and test data. Use splitting ratios of 80% training, 20% validation.

```{r}
# place your code here
```


# Modeling 

## Model Creation

Create a Support Vector Machines model for target-variable. Take all other parameters into account.

```{r}
# place your code here
```

# Predictions

Create predictions for train, and validation data. These will be probabilities.

```{r}
# place your code here
```


# Model Performance

We will compare our classifier to the baseline classifier.

## Baseline Classifier

Please calculate the baseline classifier (assignment to most frequent class). 

Hint: Now you have more than two classes, but the procedure is the same.

```{r}
# place your code here
```

## Confusion Matrix

Calculate a confusion matrix for Training Data:

```{r}
# place your code here
```

Calculate a confusion matrix for Validation Data:

```{r}
# place your code here
```

Calculate the Accuracy from the confusion matrix (for training and validation data).

```{r}
# place your code here
```

Is our classifier superior to baseline classifier?

```{r}
# put your code here
```

# Hyperparameter Tuning

Create models for a range of cost-parameters.

Hint: Inspect tune.svm() function from **e1071** package. Provide a range for parameters cost and gamma. Find the best parameter set and create a new model with these parameters.

Hint: You modified training and validation dataset. Think whether you can take all variables into account or if you need to drop some.

Calculate confusion matrices and get accuracies.

```{r}
# place your code here
```

# Acknowledgement

We thank the creators and authors of the dataset.

Creators: 

1. Hungarian Institute of Cardiology. Budapest: Andras Janosi, M.D. 
2. University Hospital, Zurich, Switzerland: William Steinbrunn, M.D. 
3. University Hospital, Basel, Switzerland: Matthias Pfisterer, M.D. 
4. V.A. Medical Center, Long Beach and Cleveland Clinic Foundation: Robert Detrano, M.D., Ph.D. 

Donor: 

David W. Aha (aha '@' ics.uci.edu) (714) 856-8779