---
title: "List Handling"
author: "Bert Gollnick"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This tutorial provides a basic introduction to lists. It is shown how you can create a list and access its elements. The key concept of applying a function to each element of a list is explained based on _purrr_ package with its _map()_ function.

# Packages

```{r}
library(purrr)
library(listviewer)
library(tidyverse)
```


# Introduction

A _list_ is a very flexible data type in R. It can collect elements of different types and lengths. It is a relative of a _dataframe_. 

A _dataframe_ is more restricted. Variables need to have unique names. Each variable is a vector and all variables have the same length.

# Creation and Accessing of Elements

A _list_ is created 
```{r}
# code here
```

An element of this list can be accessed with squared brackets. The first item of the list is accessed with:
```{r}
# code here
```

If the third member of this first item should be accessed, double brackets need to be used, follwed by the index in single brackets.
```{r}
# code here
```

You can create named lists.

```{r}
# code here
```

# View a List

Default list view is not very nice looking. For a nicer view package _listviewer_ can be used. Please make sure it is installed before loading it.
```{r}
# code here
```

# Applying a Function to each Element 

Most interesting part of this tutorial is how to apply a function to each element of the list. For this package _purrr_ needs to be loaded. Also, package _listviewer_ is used for creating a nice view. Please make sure you have installed it before loading.

## Simple Example
We will create a list with five highest mountains, their names, and their heights. Here, named lists are used with the names "name" and "height".

```{r}
Nr1 <- list(name = "Mount Everest",
	    height = 8848)
Nr2 <- list(name = "K2",
	    height = 8611)
Nr3 <- list(name = "Kangchenjunga",
	    height = 8586)
Nr4 <- list(name = "Lhotse",
	    height = 8516)
Nr5 <- list(name = "Makalu",
	    height = 8485)

# code here
```

It is not easy to extract, e.g. all names of the mountains, from this format. You would need some _for_-loop or _lapply_. Luckily, this task is very easy with _map_. As parameters we pass the list _mountains_. It also may be a column in a dataframe or a vector. Second parameter is a function. In our case the name variable should be shown.

```{r}
# code here
```

The output of _map_ is itself a list. In our example it would be more intuitive to return a character. For this there are several map-function that return a specific type (e.g. _map_int_ for integer, _map_df_ for returning a dataframe). We make use of _map_chr_ for returning a character vector.

```{r}
# code here
```

## Advanced Example

More complex functions can be used as well. In this example famous _iris_ dataset is used, which contains information on iris flowers like Sepal.Length, Sepal.Width, Petal.Length, Petal.Width, and Species. 

First, we load _iris_ data. Data is piped to the split function. (Piping concept is explained in another lecture). The results are three different dataframes. Now, _map()_ function gets active. 

```{r}
# code here
```

