---
title: "kmeans Clustering Solution"
author: "Bert Gollnick"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    code_folding: hide
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(stats))
library(factoextra)
```

# Introduction

We will work on a dataset on Pulsar candidates. This data was collected during "High Time Resolution Universe Survey (South)".

Some candidates were later confirmed Pulsars, the other could not be confirmed. This is supervised learning. Assume we don't know the true values, because this is the purpose of clustering - to find relationships that were not known before.

Here we will use the "true" values to assess how well our clustering works.

# Data Understanding

Data and description of data is found [here](https://archive.ics.uci.edu/ml/datasets/HTRU2).

HTRU2 is a data set which describes a sample of pulsar candidates collected during the High Time Resolution Universe Survey (South) [1]. 

Pulsars are a rare type of Neutron star that produce radio emission detectable here on Earth. They are of considerable scientific interest as probes of space-time, the inter-stellar medium, and states of matter (see [2] for more uses). 

As pulsars rotate, their emission beam sweeps across the sky, and when this crosses our line of sight, produces a detectable pattern of broadband radio emission. As pulsars 
rotate rapidly, this pattern repeats periodically. Thus pulsar search involves looking for periodic radio signals with large radio telescopes. 

Each pulsar produces a slightly different emission pattern, which varies slightly with each rotation (see [2] for an introduction to pulsar astrophysics to find out why). Thus a potential signal detection known as a 'candidate', is averaged over many rotations of the pulsar, as determined by the length of an observation. In the absence of additional info, each candidate could potentially describe a real pulsar. However in practice almost all detections are caused by radio frequency interference (RFI) and noise, making legitimate signals hard to find. 

Machine learning tools are now being used to automatically label pulsar candidates to facilitate rapid analysis. Classification systems in particular are being widely adopted, 
(see [4,5,6,7,8,9]) which treat the candidate data sets as binary classification problems. Here the legitimate pulsar examples are a minority positive class, and spurious examples the majority negative class. At present multi-class labels are unavailable, given the costs associated with data annotation. 

Our attributes are:

1. Mean of the integrated profile. 

2. Standard deviation of the integrated profile. 

3. Excess kurtosis of the integrated profile. 

4. Skewness of the integrated profile. 

5. Mean of the DM-SNR curve. 

6. Standard deviation of the DM-SNR curve. 

7. Excess kurtosis of the DM-SNR curve. 

8. Skewness of the DM-SNR curve. 

9. Class (target variable)

# Data Preparation

## Data Download

First we download the data.
```{r}
# if file does not exist, download it first
file_path <- "./data/HTRU2.zip"
if (!file.exists(file_path)) {
  dir.create("./data")
  url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/00372/HTRU2.zip"
  download.file(url = url, 
                destfile = file_path)
}

# unzip the file
unzip(zipfile = file_path, exdir = "./data")


```

## Data Import

The data is imported and assigned to a dataframe, called "df_raw". 1000 observations are sampled.

```{r}
set.seed(123)
file_name <- "./data/HTRU_2.csv"
df_raw <- read.csv(file = file_name, header = F) %>% 
  dplyr::sample_n(1000)
```

The right column names are assigned.

These names are used:

"Profile_mean", "Profile_stdev", "Profile_skewness", "Profile_kurtosis", "DM_mean", "DM_stdev", "DM_skewness", "DM_kurtosis", "class"

```{r}
colnames_to_set <- c("Profile_mean", "Profile_stdev", "Profile_skewness", "Profile_kurtosis", "DM_mean", "DM_stdev", "DM_skewness", "DM_kurtosis", "class")
colnames(df_raw) <- colnames_to_set
```

## Data Understanding

1. Find out if there are NA to be handled!

```{r}
summary(df_raw)
```

There are no NA's that need to be handled.

2. How many legitimate pulsars (class 1) and how many spurious results (class 0) are there?

```{r}
table(df_raw$class)
```

There are 1000 samples which are spurious.

## Data Manipulation

In clustering a distance is measured. Due to this, it is important to apply scaling to the data.

3. Create a new object called "df_scaled", which drops column "class" and scales the data.

```{r}
df_scaled <- df_raw %>% 
  dplyr::select(-class) %>% 
  scale()
```

# Model

In this chapter you will create a clustering model and evaluate it afterwards.

## Model Creation

4. Please create a clustering model based on kmeans-algorithm. With kmeans you need to assign how many clusters should be defined. Please select a reasonable number of clusters.

```{r}
mod_cluster <- kmeans(x = df_scaled, centers = 2)
```

x specifies the data. k specifies the number of clusters.

## Visualisation

5. Create a plot that shows the actual class and the cluster.

```{r}
df_raw$cluster <- mod_cluster$cluster -1
g <- ggplot(df_raw, aes(x = as.factor(class), y = as.factor(cluster), 
                        col = as.factor(class)))
g <- g + geom_jitter()
g <- g + theme_bw()
g <- g + labs(x = "Actual", y = "Cluster")
g <- g + scale_color_discrete(name = "Class Type")
g
```

```{r}

fviz_cluster(object = mod_cluster, data = df_scaled, ellipse = T, geom = "point")
```


6. Create a table that shows the count for actual class and cluster.

```{r}
table(df_raw$class, df_raw$cluster)
```




# Acknowledgement

Thanks to the author of this paper for providing information on the dataset.

Dr Robert Lyon, University of Manchester, School of Physics and Astronomy, Alan Turing Building, Manchester M13 9PL, United Kingdom, robert.lyon '@' manchester.ac.uk