---
title: "Upper Confidence Bound"
author: "Bert Gollnick"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    number_sections: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message = F, warning = F)
```

```{r}
suppressPackageStartupMessages(library(tidyverse))
library(knitr)
```


# Problem Description

Imagine you are in a casino and can choose to play at one of five Bandit. All have different distributions of returns. Which one should you choose?

# Bandit Returns

The designer of the bandits inherently designed average and variation parameters into the machines.

```{r}
set.seed(123)
bandits_design <- data_frame(Bandit = c("A", "B", "C", "D", "E"),
                             Mean = c(.97, .98, 1.1, .99, .96),
                             SD = c(.05, .15, .10, .11, .08))

bandits <- data_frame(A = rnorm(n = 1000, mean = bandits_design$Mean[bandits_design$Bandit == "A"], 
                                sd = bandits_design$SD[bandits_design$Bandit == "A"]),
                      B = rnorm(n = 1000, mean = bandits_design$Mean[bandits_design$Bandit == "B"], 
                                sd = bandits_design$SD[bandits_design$Bandit == "B"]),
                      C = rnorm(n = 1000, mean = bandits_design$Mean[bandits_design$Bandit == "C"], 
                                sd = bandits_design$SD[bandits_design$Bandit == "C"]),
                      D = rnorm(n = 1000, mean = bandits_design$Mean[bandits_design$Bandit == "D"], 
                                sd = bandits_design$SD[bandits_design$Bandit == "D"]),
                      E = rnorm(n = 1000, mean = bandits_design$Mean[bandits_design$Bandit == "E"], 
                                sd = bandits_design$SD[bandits_design$Bandit == "E"])


                                       )
bandits_gather <- bandits %>% 
  gather(key = "Bandit", value = "Return") 
bandits_gather$Return_binary <- ifelse(bandits_gather$Return < 1, 0, 2)


summary(bandits_gather)
bandit_means <- bandits_gather %>% 
  group_by(Bandit) %>% 
  summarise(Mean = mean(Return))
```

Their distributions of returns is shown in the following graph.

```{r}
g <- ggplot(bandits_gather, aes(x=Bandit, y=Return, fill = Bandit))
g <- g + geom_violin()
g <- g + geom_hline(yintercept = 1.00)
g <- g + coord_cartesian(ylim = c(.50, 1.50))
g <- g + geom_point(data=bandit_means, aes(x = Bandit, y = Mean))
g <- g + xlab("Bandit")
g <- g + xlab("Return [-]")
g <- g + ggtitle("Bandit Returns")
g <- g + theme(legend.position="bottom")

g <- g + labs(y = "Average Return [-]", x = "Round [-]", title = "Average Return and Round")
g <- g + theme_bw()
g <- g + theme(text = element_text(size=12))
g
```

And here in tabular form:

```{r}
bandits_design %>% kable()
```


We will check if our random values have the same average values.

```{r}
bandits_gather %>% 
  group_by(Bandit) %>% 
  summarise(Mean = round(mean(Return), 3)) %>% 
  kable()
```

# Exploration / Exploitation with Time

This information is only know to the designer, but not to the player. The player might perform a set of runs, e.g. 1000 runs, per machine, and find out that machine B is having the highest return. But this strategy is very costly.

We decide to play 100 rounds. In each round we decide to play on which machine based on previous knowledge.

```{r}
df_round <- data_frame(Bandit = c("A", "B", "C", "D", "E"),
                     N = 1,  # number of times the slot machine was played 
                     R = 1,  # sum of rewards
                     r = 1,  # average reward
                     delta = 1E4,  # confidence interval
                     upper_conf_bound = 1E10)  # r + delta

max_it <- 1000  # maximum iteration

df_iterations <- data_frame(n = 1:max_it,
                            SelectedBandit = NA,
                            UCB = NA,
                            r = NA,
                            delta = NA
                            )
bandit_names <- df_round$Bandit

for (n in 2:max_it) {
  # for very first iteration just start randomly
  if (n == 2) {  
    bandit_chosen_name <- "A"
    bandit_chosen_nr <- which(df_round$Bandit == bandit_chosen_name)
  }
  # increase nr of runs per selected machine
  df_round$N[bandit_chosen_nr] <- df_round$N[bandit_chosen_nr] + 1
  # get mean/sd of selected machine
  current_bandit_mean <- bandits_design$Mean[bandit_chosen_nr]
  current_bandit_sd <- bandits_design$SD[bandit_chosen_nr]
  # calculate current return
  current_return <- rbinom(n = 1, 
                              prob = current_bandit_mean/2, 
                              size = 1) * 2
  
  # calculate sum of returns
  df_round$R[bandit_chosen_nr] <- df_round$R[bandit_chosen_nr] + current_return
  # calculate average returns
  df_round$r[bandit_chosen_nr] <- df_round$R[bandit_chosen_nr] / df_round$N[bandit_chosen_nr]
  # calculate confidence interval
  df_round$delta[bandit_chosen_nr] <- sqrt(2*log(n) / df_round$N[bandit_chosen_nr])
  # calculate upper confidence bound
  df_round$upper_conf_bound[bandit_chosen_nr] <- df_round$r[bandit_chosen_nr] + df_round$delta[bandit_chosen_nr]

  # store selection in df_iteration
  df_iterations$SelectedBandit[n] <- bandit_chosen_name
  df_iterations$UCB[n] <- df_round$upper_conf_bound[bandit_chosen_nr]
  df_iterations$r[n] <- df_round$r[bandit_chosen_nr]
  df_iterations$delta[n] <- df_round$delta[bandit_chosen_nr]
  
  # define bandit for next run
  bandit_chosen_nr <- which.max(df_round$upper_conf_bound)
  bandit_chosen_name <- bandits_design$Bandit[bandit_chosen_nr]
}  
```

```{r}
g <- ggplot(df_iterations %>% dplyr::filter(!is.na(SelectedBandit)), aes(x = n,
                               y = r, 
                               col = SelectedBandit))
g <- g + geom_point()
g <- g + theme_bw()
g <- g + theme(text = element_text(size=12))
g <- g + theme(legend.position="bottom")
g <- g + labs(y = "Average Return [-]", x = "Round [-]", title = "Average Return and Round")
g
```

# Exploration / Exploitation at specific Timesteps

The graph shows for the current round the confidence bounds for all bandits. The red bar represents the bandit with the highest upper confidence bound, that will be selected for the next round.


```{r}
g <- ggplot(df_round, aes(x = Bandit, 
                          y = r))
g <- g + geom_point()
g <- g + geom_errorbar(aes(ymin = r - delta,
                           ymax = r + delta))
g <- g + geom_point(data = df_round[which.max(df_round$upper_conf_bound), ], col = "red")
g <- g + geom_errorbar(data = df_round[which.max(df_round$upper_conf_bound), ], col = "red",
                       aes(ymin = r - delta,
                           ymax = r + delta))
g <- g + labs(x = "Bandit", y = "Confidence Bound [-]", title = "Confidence Bound")
g <- g + theme_bw()
g <- g + theme(text = element_text(size=12))
g <- g + theme(legend.position="bottom")
g
```

