---
title: "Using Pretrained Models"
author: "Bert Gollnick"
output: 
  html_document:
    toc: true
    number_sections: true
    code_folding: hide
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

# Packages

```{r}
library(tidyverse)  # data plotting and manipulation
library(keras)  # deep learning
library(jpeg)  # image processing
```

# Data Preparation

## Loading a pretrained model

At first we load the trained model.

```{r}
vgg19 <- application_vgg19(weights = "imagenet",
                           include_top = T)
vgg19
```

## Prepare an image

We will use my profile image and an image of an elephant (you can exchange it and use an own image).

This function is used to plot the image.

```{r}
plot_jpeg = function(path, add=FALSE)
{
  jpg = jpeg::readJPEG(path, native=T) # read the file
  res = dim(jpg)[2:1] # get the resolution, [x, y]
  if (!add) # initialize an empty plot area if add==FALSE
    plot(1,1,xlim=c(1,res[1]),ylim=c(1,res[2]),asp=1,type='n',xaxs='i',yaxs='i',xaxt='n',yaxt='n',xlab='',ylab='',bty='n')
  rasterImage(jpg,1,1,res[1],res[2])
}
```

Now we define which photo to use.

```{r}
image_file <- "./data/elephant.jpg"

plot_jpeg(image_file)
```

Now we need to rescale it and import it to an array. We can make use of keras functions for this purpose.

```{r}
img <- image_load(path = image_file, target_size = c(224, 224)) %>% 
  image_to_array() %>%  # cast it to an array
  array_reshape(x = ., dim = c(1, dim(.))) %>%   # add an additional dimension
  imagenet_preprocess_input(mode = "tf")

```

# Model Predictions

```{r}
predictions <- vgg19 %>% predict(x = img) 
pos_max <- which.max(predictions)
```

## Get Class Labels

```{r}
class_labels <- read.table("./data/imagenet1000_clsidx_to_labels.txt", sep = ":") %>% 
  select(-V1) %>% 
  rename(label = V2)
class_labels$label[pos_max]
```

We used an image of an indian elephant and received the right prediction.

Now let's see what other predictions were.

```{r}
class_labels$preds <- predictions[1, ]

class_labels %>% 
  arrange(desc(preds)) %>% 
  top_n(5) %>% 
  mutate(label = factor(x = label, label)) %>% 
  ggplot(data = ., aes(x = label,
                       y = preds)) + 
  geom_col() + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

