---
title: 'Web Scraping'
author: "Bert Gollnick"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
suppressPackageStartupMessages(library(knitr))
```

# Introduction

Web scraping is used to extract data from websites. This is incredible easy with R. In this tutorial you will learn how to extract a table from a complex site. In the Wikipedia article (Wind power by country)[https://en.wikipedia.org/wiki/Wind_power_by_country] you find a table with installed wind power capacity, embedded in a complex site with text, graphs and other tables. How can you extract information of this specific table? 

* Objective: Learn how to extract components of websites
* Requirements: Google Chrome, R Basics

# Preparation

First, you need to install and load the package _rvest_.

```{r get_package}
# install package once
#install.packages("rvest")

# load library rvest
library("rvest")
```

Now you can define the url.
```{r url}
url <- "https://en.wikipedia.org/wiki/Wind_power_by_country"
```

To get the information of the specific table you need to get _xpath_ information. _xpath_ can be used to navigate in XML documents. We will use it to identify the table of interest. If you have installed Google Chrome it is easy. (If you are using a different browser, there are many addins for this purpose.)

If you in the browser and on the website, just click CTRL + SHIFT + I or right-click and Inspect. On the right side of your screen a new windows appears called DevTools. If you hover over the items the objects on the website are highlighted. Now you browse until you find the table highlighted. 

Now you only have to copy the xpath string by _right-clicking_ --> _Copy_ --> _Copy XPath_. This is used within R for accessing the table. This reference can be used in html_node(). At first the complete



```{r}
url <- 'https://en.wikipedia.org/wiki/Wind_power_by_country'
page <- read_html(url)

# Find the table element containing wind generation data 
table <- page %>% 
  html_nodes("table.wikitable.sortable") %>%
  .[[1]]

# Get the table header
header <- table %>%
  html_nodes("th") %>%
  html_text()

# Get all rows of data from the table
data <- table %>%
  html_nodes("tr") %>%
  .[2:length(.)] %>%
  html_nodes("td") %>%
  html_text()

# Reshape data into a data frame
df <- matrix(data = data, ncol=length(data)/6, nrow = 6) %>% data.frame() %>% t() %>% data.frame()
  
# Set column names of the data frame
colnames(df) <- header

# Print the data frame
print(df)
```

