---
title: "Data Filtering: Solution"
author: "Bert"
date: "18 März 2019"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F)
```

```{r}
library(tidyverse)
library(babynames)
```

In this exercise you will work extensively on dplyr filtering and selection.

# Data Preparation

We will work with dataset "babynames" from the package with the same name.

```{r}
data("babynames")
babynames %>% glimpse
```

# Filtering

1. Filter for year 2015.

2. Filter for all names that have a capital letter "B". Help: take a look at "str_detect()".

3. Filter for male names.

4. Take only names into account which have more than 100 observations.

5. Show the Top 10 in decreasing order.

6. Keep only the columns "name" and "prop"

```{r}
babynames %>% 
  filter (year == 2015) %>% 
  filter(str_detect(name, "B")) %>% 
  filter(sex == "M") %>% 
  filter (n >= 100) %>% 
  arrange(desc(n)) %>% 
  top_n(10) %>% 
  select(name, prop)
```

