---
title: "Aggregating Results: Exercise"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F)
```

```{r}
library(tidyverse)
```


# Data Preparation

# Exercise

- Load the dataset "txhousing", which holds information on Texas housing market.

```{r}
data("txhousing")
txhousing %>% head
```

- Calculate the average sales for each city. During the calculation of mean values the result is coerced to NA, if there is a single NA present. Disable this functionality and change it so that missing values are excluded.

```{r}
txhousing %>% 
  group_by(city) %>% 
  summarise(mean_sales = mean(sales, na.rm = T))
```

- Calculate the average sales for city and year and save it in object "tx_city_year".

```{r}
tx_city_year <- txhousing %>% 
  group_by(city, year) %>% 
  summarise(mean_sales = mean(sales, na.rm = T))
```

```{r}
ggplot(tx_city_year, aes(x = year, y = mean_sales, col = city)) + 
  geom_line() + 
  labs(x = "Year", y = "Mean Sales", title = "Average Sales per Year and City")
```

