---
title: "Joining Datasets"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message = F, warning = F)
```

# Data Preparation

```{r}
library(tidyverse)
```


We create two dataframes that we will join then together to one dataset.

```{r}
a <- tibble(index = 1:5, 
            value = rnorm(5))

b <- tibble(index = 3:7, 
            value = rnorm(5))
a
b
```

# Left Join

Left join takes the LHS dataset (first dataframe) and keeps it unchanged. Based on a common index-column, the values of RHS (second dataframe) are bound at the locations for which index locations are available.

```{r}
# code here
```

# Right Join

Right join takes the RHS dataset (second dataframe) and keeps it unchanged. Based on a common index-column, the values of LHS (first dataframe) are bound at the locations for which index locations are available.

```{r}
# code here
```

# Full Join


Full join takes the LHS and RHS datasets and keeps them unchanged. Based on a common index-column, the values are added at the locations for which index locations are available. For positions where no information is available NA alues are added.


```{r}
# code here
```
