---
title: "Introduction to Plotly"
author: "Bert Gollnick"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message = F, warning = F)
```

```{r}
library(tidyverse)
library(plotly)
```

# Introduction

We will use data from World Happiness Report. The World Happiness Report is a landmark survey of the state of global happiness.

# Data Preparation

The data can be downloaded from [here](https://www.kaggle.com/unsdsn/world-happiness/).

```{r}
happiness_2015 <- read_csv("./data/2015.csv") %>% 
  mutate(year = 2015)
happiness_2016 <- read_csv("./data/2016.csv") %>% 
  mutate(year = 2016)
happiness_2017 <- read_csv("./data/2017.csv") %>% 
  mutate(year = 2017)

happiness <- happiness_2015 %>% 
  bind_rows(happiness_2016) %>% 
  bind_rows(happiness_2017) %>% 
  select(Country, `Happiness Score`, `Economy (GDP per Capita)`, `Health (Life Expectancy)`, Freedom, year)
happiness %>% glimpse()
```

```{r}
happiness_mod <- happiness %>% 
  rename(happiness_score = `Happiness Score`) %>% 
  rename(economy_gdp = `Economy (GDP per Capita)`) %>% 
  rename(health_life_exp = `Health (Life Expectancy)`)
```



# Visualisation

## Complex Plot

We create a complex visualisation (5 dimensional), which has these properties:

- x axis refers to GDP (an economical value)

- y axis refers to life expectancy (a health value)

- size refers to freedom

- color refers to happiness score

- text refers to country

- slider refers to the year

```{r}
# code here
```


## Using Subplots

```{r}
# code here
```

# More Information

More information on plotly can be found on the [R Plotly Homepage](https://plotly-r.com).