---
title: "Outlier Detection: Solution"
author: "Bert Gollnick"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(dbscan)
```


# Introduction

The data set you work with includes shapes, that are easily to distinguish for a human. Your task is to write an algorithm to detect the shapes with an algorithm.

```{r}
library(factoextra)
data("multishapes")
```

```{r}
g <- ggplot(multishapes, aes(x, y, col = as.factor(shape)))
g <- g + geom_point()
g
```

Hint: make use of dbcan for this exercise.

# dbscan

Create a matrix object of "multishapes".

Create a kNNdistplot for 5 points.

```{r}
multi_mat <- as.matrix(multishapes)
kNNdistplot(multi_mat, k = 5)
```

Use the plot to come up with a suitable epsilon parameter. 

Then use dbscan to detect the clusters with this algorithm.

```{r}
eps <- 0.2
min_points <- 5

db <- dbscan(x = multi_mat, eps = eps, minPts = min_points)
```


Assign a new column in "multishapes" based on the clusters that were detected by dbscan.

```{r}
multishapes$dbscan_cluster <- db$cluster
```

Plot the result.

```{r}
g <- ggplot(multishapes, aes(x, y, col = as.factor(dbscan_cluster)))
g <- g + geom_point()
g
```

Did your parameters lead to the right clusters? If not, change parameters and run it again.

