---
title: "Stringr and Regular Expressions"
author: "Bert Gollnick"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

# Introduction

A good introduction on stringr and regular expressions can be found [here](https://stringr.tidyverse.org/articles/regular-expressions.html).

You also can make use of the (cheatsheet)[https://www.rstudio.com/wp-content/uploads/2016/09/RegExCheatsheet.pdf].

# Packages

```{r}
library(stringr)  # stringr is part of tidyverse
```

Stringr is part of the tidyverse, so we don't need to load it separately.

```{r}
string <- c('R for Data Handling, Data Science, and Machine Learning 2020',
            'Python for Machine Learning',
            'Julia for Statistics'
            )
string
```

# Detect Patterns

We can detect patterns with grep or grepl (base R) or with functions from string r package, like str_detect.

```{r}
pattern <- "ata"

# return logical for each element
str_detect(string = string, pattern = pattern)  # similar to grepl
grepl(x = string, pattern = pattern)

grep(x = string, pattern = pattern)

# return value, that includes the string
grep(x = string, pattern = pattern, value = T)
```

# Locate Patterns within string

With **str_locate** we can find the location of a pattern within the string, or with **str_locate_all** for all locations in the string.

```{r}
str_locate(string = string, 
           pattern = pattern)  # finds start and end position of FIRST matches
str_locate_all(string = string, pattern = pattern)  # finds start and end position of all matches
```

# Replace Patterns

We can replace patterns with **str_replace**. A specific pattern is replaced by another pattern.

```{r}
str_replace(pattern = "Data", string = string, replacement = "DATA")  # similar to sub
str_replace_all(pattern = "Data", string = string, replacement = "DATA")  # similar to gsub
```

# Split a String

The string can be splitted at specific delimiters.

```{r}
pattern <- " "

str_split(string = string[1], pattern = pattern, simplify = T)
```

# Extract Patterns

Patterns can be extracted with **str_extract**.

```{r}
pattern <- "Learning"
str_extract_all(string = string, pattern = pattern, simplify = T)
str_match_all(string = string, pattern = pattern)
```

# Regular Expressions

So far we used normal text as patterns. Now we will make it more generic.

## Character Classes

You can define a pattern for digits [[:digit:]] or non-digits "\\D"

```{r}
# detect digits
pattern <- "[[:digit:]]"
str_detect(string = string, pattern = pattern)

# detect non-digits
pattern <- "\\D"
str_detect(string = string, pattern = pattern)

```

Similar you can search for lower case [[:lower:]] and upper case [[:upper:]].

```{r}
pattern <- "[[:lower:]]"

str_detect(string = string, pattern = pattern)
```

It does check if there are lower case characters and not if all characters are lower-case.

Other useful character classes are 

- . any character

- | or, like in (a|b)

- [...]  permitted characters, e.g. [abcd]

- [a-d]  character ranges


## Anchors

Normally the patterns are searched in the complete string. If you want to restrict the pattern to be found only at the beginning or at the end.

- ^  matches at the beginning

- $  matches at the end

Search for words starting with "H".

```{r}
pattern <- "^R"
grep(x = string, pattern = pattern)
```

Search for words with an "a" at the second position.

```{r}
pattern <- "^.y"
grep(x = string, pattern = pattern)

```

## Quantifiers

The matching can occure a number of times.

- "?" for zero or one matching

- "+" for one or more matchings

- "*" for zero or more matchings

We want to search for a number with at least 3 digits.

- "{n}"  matches n times

- "{n,m}"  matches between n and m times

```{r}
string <- c("1", "10", "100", "1000")
pattern <- "\\d{3}"
str_extract(string = string, pattern = pattern)
```



## Look Arounds

With look arounds you can look before or after the matching pattern.

- (?=...)  positive look-ahead (it matches when the dots match the input)

- (?!...)  negative look-ahead

- (?<=...) positive look-behind

Now we have a more complex example.

```{r}
string <- c("100 course", "2 courses", "1 instructor")
```

We want to detect the locations of "1 course" or "2 courses". For this, the word course or courses needs to be found together with a precending space and a number.

```{r}
pattern <- "\\d*(?= course?)"
str_detect(string = string,
            pattern = pattern)
```


