---
title: "Univariate Linear Regression"
author: "Bert Gollnick"
date: "November 23, 2018"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Understanding

We will work with a dataset, that is provided by **dplyr** package. Based on star wars characters we will predict person weight based on its height.

What is independent and dependent variable? Well, the height is more or less defined in our genes, but our weight can be influenced. So I use height as independent variable ad mass as dependent variable.

## Packages

We load required packages.

```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
```

## Data Import

We load the data.

```{r}
data("starwars")
set.seed(123)
```

## Visualisation

We write a small function for visualising the data.

```{r}
plot_height_mass <- function(df) {
  g <- ggplot(df, aes(x = height, y = mass))
  g <- g + geom_point()
  g <- g + geom_smooth(method="lm", se=F)
  g
}

plot_height_mass(starwars)
```

# Data Preparation

There is a clear outlier we need to handle.

```{r}
starwars %>% 
  filter(mass > 1000) %>% 
  select (name, height, mass)
```


```{r}
starwars_filt <- starwars %>% 
  filter (mass < 1000)

plot_height_mass(starwars_filt)
```

# Modeling

Model is created with *lm()*.

```{r}
model <- lm(data = starwars_filt, 
            formula = mass ~ height)
```

We take a look at the summary.

```{r}
model_summary <- model %>% 
  summary()
model_summary
```

R**2 is calculated.

```{r}
model_summary$r.squared
```

# Predictions

```{r}
starwars_filt$mass_predictions <- predict(object = model, 
                       newdata = starwars_filt)
```

```{r}
g <- plot_height_mass(starwars_filt)
g <- g + geom_point(data = starwars_filt, 
                    aes(x = height, 
                        y = mass_predictions),
                    col = "red", alpha = .3)
g
```

We see that our predictions exactly match the smoothed line from ggplot.

But how good do they match our actual data.

```{r}
g <- ggplot(data = starwars_filt,
            aes(x = mass,
                y = mass_predictions,
                label = name)
            )
g <- g + geom_point()
g <- g + geom_text(size = 3)
g <- g + geom_abline(slope = 1, 
                     intercept = 0,
                     col = "blue")
g <- g + labs(x = "Actual Mass [lb]",
              y = "Predicted Mass [lb]",
              title = "Mass Correlation")
g
```
