---
title: "Univariate Regression: Solution"
author: "Bert Gollnick"
date: "October 25, 2018"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    code_folding: hide
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library("lava"))  # includes data
```

# Data Understanding

In this tutorial you will take a look at measurements of Hubble (the telescope). Besides taking beautiful pictures, it measured speed and distance of Super-Novae. Similar data was used in 1929 by Hubble (the person) and he found out that there is a linear relationship.

He discovered that galaxies appear to move away. This can be visualised with red-shift of spectral lines. This observation was the first indication that the universe expands.

You will create a linear model based on observations and create predictions.

# Data Preparation

## Data Import

You import the data with **data()**.

```{r}
data("hubble")
```

Show the first observations of the dataframe.

```{r}
hubble %>% head()
```


It holds Velocity (v) and distance (D) measures of 36 Type Ia super-novae from the Hubble Space Telescope.

The data is visualised.

```{r}
g <- ggplot(hubble, aes(v, D))
g <- g + geom_point()
g <- g + geom_smooth(method = "lm", se = F)
g
```

We can clearly observe the linear relationship of the data.

# Modeling

You will create a linear model. Dependent variable is "D", independent variable is "v".

What are the values of parameters for intercept and slope for "v"?

```{r}
fit_lm <- lm(formula = D ~ v, 
             data = hubble)
fit_lm
```

Calculate adjusted R squared.

```{r}
fit_lm_summary <- summary(fit_lm)
fit_lm_summary$adj.r.squared
```

Is there a high correlation between D and v?

```{r}
# Yes, 97.8% is an extremely high correlation coefficient.
```

# Predictions

A new dataframe is created, that covers a broad range of speeds.

```{r}
new_data <- data.frame(v = seq(0, 30000, 1000))
```

Add to the dataframe "new_data" a column "D_pred", that hold predictions of "D", based on the model you created.

```{r}
new_data$D_pred <- predict(object = fit_lm, newdata = new_data)
```

Now we can compare actual data and plot our linear regression line on top.

```{r}
g <- ggplot(new_data, aes(v, D_pred))
g <- g + geom_line(color = "blue", size = 2)
g <- g + geom_point(data = hubble, aes(v, D))
g <- g + theme_bw()
g <- g + xlab ("Speed v [km/s]")
g <- g + ylab ("Distance D [MPc]")
g <- g + ggtitle("Hubble Super-Novae Type Ia Measurements and Predictions")
g
```

Our linear regression line perfectly matches the one shown in the previous plot. So ggplot uses the same calculation for **geom_smooth()**.

# Bonus Question: Hubble Constant

Hubble constant (H0) is one of the most important values in cosmology. 

$$H_0 = \frac{v}{D}$$

```{r}
hubble$H_0 <- hubble$v / hubble$D
median(hubble$H_0)
```

You can compare this to most recent observed values, as shown in this [Wikipedia article](https://en.wikipedia.org/wiki/Hubble%27s_law).

# Conclusion

You learned to create a linear (and univariate) model, which is the most basic model you can create. But if the basics are understood, more complicated models can be created with little effort.

# Acknowledgement

We thank the author of the data for providing it:

Freedman, W. L., et al. 2001, AstroPhysicalJournal, 553, 47.