---
title: "Polynomial Regression"
author: "Bert Gollnick"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Understanding and -preparation

We create an artificial dataset from scratch to show the principle.

# Packages

We load required packages.

```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
```

## Data Creation

This time we won't load data, but instead create some data to play with.

For a sequence of **x** we create noisy data with a follow a third order polynomial trend.

```{r}
sample_data <- tibble(x = seq(-20, 40, 0.5),
                      y = 50 + 0.25 * (x-5)**3,
                      y_noise = y + rnorm(n = length(y),
                                          mean = 100, 
                                          sd = 500))
```

## Visualisation

```{r}
g <- ggplot(sample_data, aes(x = x,
                             y = y_noise))
g <- g + geom_point()
# g <- g + geom_smooth(method = "lm", se = F)
g <- g + geom_line(aes(y = y), col = "red", size = 1)
g 
```

# Model

# Linear Model

A linear trend is clearly a poor choice. We expect that a third order polynomial is a good approximation. 

```{r}
# add code here
```

```{r}
# add code here
```

Adjusted R-squared is 0.76. This is an example of underfitting. Our model is not complex enough to cover the complexity of the data.

# Quadratic Model

```{r}
# add code here
```


# Polynomial (3rd order) Model 

So we create a model that estimates the parameters.

You might do it this way.

```{r}
# add code here
```

To save some coding, you might use **poly()** function.

```{r}
# add code here
```

We can analyse our parameter estimates and the summary of the model.

```{r}
# add code here
```

R-squared is now 0.977 and much higher compared to a linear model.

We also can check residuals and model fitted values to investigate if there is a pattern in the data. If so, that would mean, our model did not fit well.

```{r}
# add code here
```

Many points fit to 0, which is not surprising (take a look at the plot before). Besides this, there is no clear pattern.

# Predictions

We predict data based on the model. 

Hint: Usually it should be avoided to predict values based on a model, that was trained on the same data. We will introduce resampling techniques very soon.

```{r}
# add code here
```

We check the prediction by comparing it to our actual data.

```{r}
g <- ggplot(sample_data, aes(x, y_pred))
g <- g +geom_point(aes(y = y_noise))
g <- g + geom_line(col = "green", size = 2)
g <- g + geom_line(data = sample_data, aes(x, y), col = "red")
g
```

# Model Performance

```{r}
# add code here
```

R squared measures proportion of variation of dependent variable **y**, that is explained by independent variables **x** for a linear model.

Adjusted R squared calculates the metric, if more than one independent variable is used.

If you have more than one independent variable - use adjusted R-squared. For only one independent variable, both are calculating the same and are interchangable.



