---
title: "Multivariate Regression Exercise"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F)
```

# Data Understanding

You will work on airfoil self noise prediction. Data comes from NASA airfoil testing. Airfoil blade sections were tested with regards to sound emission.

These are the independent variables:

- Frequency, in Hertzs. 

- Angle of attack, in degrees. 

- Chord length, in meters. 

- Free-stream velocity, in meters per second. 

- Suction side displacement thickness, in meters. 

The target variable is 

- Scaled sound pressure level, in decibels. 

```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(corrplot))
suppressPackageStartupMessages(library(caret))
```


# Data Preparation

```{r}
# if file does not exist, download it first
file_path <- "./data/airfoil_noise.txt"
if (!file.exists(file_path)) {
  dir.create("./data")
  url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/00291/airfoil_self_noise.dat"
  download.file(url = url, 
                destfile = file_path)
}
```

Import the data and assign it to object *airfoil_noise*:

```{r}
# add code here
```


Assign the right column names. Please use these names:

frequency, aoa, chord_length, free_stream_vel, ss_displacement_thickness, spl)

```{r}
colnames(airfoil_noise) <- c("frequency", "aoa", "chord_length","free_stream_vel", "ss_displacement_thickness", "spl")
```

## Exploratory data analysis

Create a correlation plot for the data to understand which variables are most valuable to predict sound pressure level.

```{r}
# add code here
```

Which two variables correlate most to target variable?

```{r}
# add code here
```

# Model1

## Model Creation

Create a multivariate linear regression model based on all independent variables.

```{r}
# add code here
```

## Model Performance

Analyse model performance and get R**2 value:

```{r}
# create predictions

# add code here

# perform model performance

# add code here
```

Does the model overfit?

Create a visualisation of predicted and actual values.

```{r}
g <- ggplot(airfoil_noise, aes(x = spl, y = spl_pred))
g <- g + geom_point()
g <- g + geom_smooth(method = "lm", se = F)
g <- g + labs(x = "Actual", y = "Predicted", title = "Correlation Plot")
g <- g + theme_bw()
g
```


# Model2

## Model Creation

Create a second model with only three independent variables. Which ones do you choose? 

```{r}
# add code here
```

## Model Performance

Analyse model performance and get R**2 value:

```{r}
# create predictions

# add code here

# perform model performance

# add code here
```

Did the model performance increase?

```{r}
# add code here
```


# Acknowledgement

This dataset comes from UCI Machine Learning Repository. Here is more information on donor and creators of the data.

Donor: 
Dr Roberto Lopez 
robertolopez '@' intelnics.com 
Intelnics 

Creators: 
Thomas F. Brooks, D. Stuart Pope and Michael A. Marcolini 
NASA